Psychological Needs During Intergroup Contact: Three Experience Sampling Studies

### 2022-03-15 TitlePage.tex
Updates from Overleaf

We have no known conflict of interest to declare. The authors received no specific funding for this work. Materials and  software is available at \url{https://janniscodes.github.io/intergroup-contact-needs/}  \citep{KreienkampMasked2022}. Protocols, materials, data, and code are available at \url{https://osf.io/pr9zs/?view_only=1ea47bb646694632a764dead807ef970} \citep{KreienkampMasked2022a}.

>>>>

We have no known conflict of interest to declare. The authors received no specific funding for this work. Materials and  software is available at \url{https://janniscodes.github.io/intergroup-contact-needs/}  \citep{Kreienkamp2022}. Protocols, materials, data, and code are available at \url{https://osf.io/pr9zs/?view_only=208a53a1f0ff48dda1c17357328fa578} \citep{Kreienkamp2022a}. The preregistration of Study 3 can be accessed as part of our Open Science Framework repository \citep{Kreienkamp2021f}.

### 2022-03-15 manuscript.tex
Updates from Overleaf

We have no known conflict of interest to declare. The authors received no specific funding for this work. Materials and  software is available at \url{https://janniscodes.github.io/intergroup-contact-needs/}  \citep{KreienkampMasked2022}. Protocols, materials, data, and code are available at \url{https://osf.io/pr9zs/?view_only=1ea47bb646694632a764dead807ef970} \citep{KreienkampMasked2022a}.

@@

% Abstract
\abstract{
One challenge of modern intergroup contact research has been the question of when and why an interaction is perceived as positive and leads to better intergroup relations. This question becomes especially relevant when considering that negative intergroup contacts are common during everyday life. To understand positive intergroup interactions, we propose to consider situational motivations and psychological needs during everyday intergroup contact. We conducted three extensive longitudinal studies with recent migrants, to capture their interactions with the majority outgroup (total N of measurements = 10,297). Across the three studies, we find inconsistent results for a traditional test of the contact hypothesis with experience sampling data. However, when using multilevel models to test the contact hypothesis, we find that individual outgroup interactions and perceived interaction quality have a strong influence on positive outgroup attitudes. We find that the fulfillment of psychological needs offers a strong predictor of positive attitudes after interactions and that these positive effects emerged via perceived interaction quality. The situational needs remained the core predictor of outgroup attitudes even after controlling for Allport’s (\citeyear{Allport1954b}) contact conditions or more general fundamental needs. As one of the first studies to test intergroup contact theory using extensive longitudinal data, we offer insight into the mechanisms of positive intergroup contact during real-life interactions and find situational motivations to be a key building block of understanding and addressing positive intergroup interactions.

@@

\section{Need Mechanism in Intergroup Contact}
% Motivational mechanism:
% interaction quality is important
% past research has either focused on conditions or cognitive-affective processes
% neither are necessary or explain why an interaction is positive
Several meta-analytic reviews found that positive intergroup contact reduces prejudice and increases positive attitudes in experimental and cross-sectional studies \citep[][]{Tropp2005, Pettigrew2006, Davies2011}, as well as in intergroup contact interventions outside the lab \citep[][]{Beelmann2014, Lemmer2015}. It is widely accepted that equal status, common goals, collaboration, and structural support during the interaction form the optimal conditions for such positive contact effects \citep[Allport's Optimal Contact conditions;][]{Allport1954b, Pettigrew1969}. However, a major meta-analysis of these contact conditions showed that contact resulted in more positive intergroup relations even when Allport's conditions were not met \citep[][]{Pettigrew2006}. It, thus, remains unclear when and why exactly an interaction is perceived as positive and what factors (beyond Allport's conditions) explain positive contact effects\footnote{It should be noted here that following Allport's original conditions, several additional conditions of optimal contact were proposed \citep[for a critical discussion see][]{Pettigrew1986}. Similarly, since the early 21\textsuperscript{st} century many psychological processes during intergroup contact were examined \citep[e.g. see,][]{Paolini2021}. Among others, researchers have for example explored different forms of social categorizations \citep[][]{Pettigrew1998}, the salience of social categories \citep[][]{Brown2005}, intimacy \citep[e.g.,][]{Marinucci2021} and attachment \citep[e.g.,][]{Tropp2021}, threat and intergroup anxiety \citep[e.g.,][]{Stephan2008, Paolini2004}, and to a lesser extent knowledge about the other group \citep[][]{Pettigrew2008c}. However, all these advances share the underlying criticism that they are not geared towards understanding when and why an interaction is perceived as positive.}. We propose that we may look to the literature on motivation and psychological needs to understand when and why exactly an interaction is perceived as positive.

@@

\section{Real-Life Process Data}
While we have argued that a need fulfillment mechanism is relevant to intergroup contact generally, its flexible and broad applicability might be ideally suited to address the pressing issue of understanding natural intergroup contacts outside the lab. Investigations of such 'real-life interactions' often suffer from the difficulty that past intergroup contact research has either focused on the mechanisms of individual interactions in artificial lab studies or has focused on longer-term recall self-reports of natural interactions \citep[e.g.,][]{Pettigrew2006}. Even with extended intervention studies, the most fine-grained data available is usually limited to pre-post-control designs. Such empirical practices stand, however, in stark contrast to many of the theoretical advances that have focused on the dynamic nature of intergroup relations \citep[e.g.,][]{Pettigrew1998}, as well as the core idea of the contact hypothesis, which was focused on the daily interactions of people \citep[see][]{Allport1954b}. As a result, prominent figures within the intergroup contact field have called for studies that collect longitudinal \citep[][]{Pettigrew1998, Pettigrew2008, Pettigrew2008b, Pettigrew2011} and real-life experience-sampling data outside the lab \citep[][]{MacInnis2015, McKeown2017, Dixon2005}. Such data would be able to capture real-life interactions that include interaction-specific mechanism information close the actual experience\footnote{Additionally, such experience-sampling data can be collected close to the intergroup interactions and would, thus, largely mitigate recall biases. Moreover, because data is nested within participants, experience-sampling data often allows to capture large amounts of high-quality data with relatively few participants.}.

@@

To test our main hypotheses, we collected three independent sets of extensive longitudinal data (Studies 1–3). The full surveys are available in Supplementary Materials A as well as in our open science repository (including a complete codebook, see OSF-REFERENCE). Our most comprehensive study (Study 3) was fully preregistered  \citep[available at][]{Kreienkamp2021f}. The fully annotated analyses are available in Supplementary Material B. As is common with multilevel analyses, we use a hierarchical modeling approach and always report the final model in-text (for the full modeling process see Supplementary Material B).

@@

\section{Discussion}
% aims re-iterated
The aim of this project was in essence threefold. We (1) tested the basic predictions of the intergroup contact hypothesis and Allport's optimal conditions in real-life extensive longitudinal data. As a flexible alternative, we (2) proposed that the fulfillment of situational psychological needs meaningfully predicts positive outgroup attitudes. Finally, we (3) assessed whether the fulfillment of psychological needs (and Allport's conditions) have their effect (partially) through perceived interaction quality.

% Contact hypothesis in ESM data [inconsistent in aggregate but consistent in ML analysis]
When considering the results of the three studies jointly, we found mixed results for the basic intergroup contact hypothesis. More specifically, we observe that separating between-participant and within-participant effects is crucial. For between-subjects effects (using aggregated reports), neither the number of interactions, nor the average interaction quality were significant predictors of positive outgroup attitudes across all studies (also see Figure \ref{fig:ContactHypothesis}a). The aggregate findings are surprising because most past cross-sectional studies that have investigated the effect of interaction frequency have also considered an aggregate measurement (albeit not one from actual daily reports). The absence of this aggregate effect could underline the fact that cross-sectional retrospective data might be misleading because (1) it presents a mixture of within and between-subjects effects \citep[][]{Hamaker2020}, or (2) suffers from recall biases (e.g., where times with no outgroup interactions are undervalued by participants during the retrospective evaluations).\footnote{It should be noted that such inconsistencies with past research might in part be a data artefact (e.g., because most people reported substantially more measurements during which they did not have an outgroup interaction). However, the multilevel models we employed should in principle be resistant to such imbalances.} In support of this possibility, once we used multilevel modeling to separate effects into a between-subjects and within-subjects part, we found that when having an interaction with an outgroup member (vs. not having an outgroup interaction) is a strong predictor of outgroup attitudes across all three studies (even when taking other non-outgroup interactions into account; also see Figure \ref{fig:ContactHypothesis}b). We thus find that when we aggregate interaction frequencies and interaction quality from our intensive longitudinal data, we find inconsistent results from the past literature, however within participants, intergroup contacts are significantly related with positive outgroup attitudes.

@@

\subsection{Limitations}
% (1) Sample = minority and (voluntary) migrants, (2) Methodology = short scales (reliability) and no lagged effects, (3) Core Need Concept = non-specific
Among the limitations of our studies we should consider the following. One limitation is our choice of samples. While we believe that a need–based mechanism should be relevant to any inter-group contact, our sample has focused on a minority- and (voluntary) migrant perspective. Without additional evidence it thus remains difficult to judge whether motivational effects will generalize to other migrant groups (e.g., forced migrants), other intergroup contexts (e.g., gender- or sexual orientation groups), and to majority groups in their outgroup attitudes. We know of no research that for other contexts psychological needs would not be relevant (once existential needs are met) but future research may extend our findings to other contexts to build an even broader understanding of psychological needs in intergroup contacts. A second limitation lies in our methodology. While extensive longitudinal data is close to the real-life events, this method comes at the expense of longer and more robust scales. Long and repetitive scales are often not feasible in extensive longitudinal methods because of the increased burden to the participants. To circumvent this shortcoming, we have ensured that the measures we used were, whenever possible, based on past validations. However, the circumstance remains that extensive longitudinal data often does not allow the same scrutiny of measurement reliability as single–shot cross–section data sets. An additional methodological question lies in the unexplored potential of the longitudinal aspects of our data. For our research questions, we have focused on contemporaneous effects within the data set, yet future investigations should seek to extend the mechanism to developmental trajectories within and between participants. And finally, our conceptualization of core situational needs has been focused on the most essential test of a motivational mechanism. This has come at the expense of specificity in the situational needs (i.e., we have not explored which exact needs people had during specific interactions). With the adaptive measurement we have proposed such an investigation would be possible (e.g., by clustering or classifying the free text entries) but would not have been relevant to our theory–focused research question. Nonetheless, future research may explore which exact needs are most important in different intergroup contexts.

\subsection{Implications}
% ESM data
Given these limitations, we can draw a number of implications for other researchers and practitioners from our results. A first implication concerns the feasibility and usefulness of extensive longitudinal data for intergroup contact research and the broader field of social psychology. While setting up an extensive longitudinal study is in no manner easy, given the amount and quality of the data, we believe the efforts to be similar to a sizable cross-section data collection (i.e., for a longitudinal or a high-quality cross-sectional data set with 10,000 data-points). Given the comparable effort, this opens up the possibility to explore research questions that focus on real-life phenomena outside the lab or focus on phenomena that depend on changes and influences over time. In the context of intergroup contact research, we were among the first to answer calls to test intergroup contact mechanisms using on extended real-life data \citep[e.g.,][]{Pettigrew2011, MacInnis2015}. In doing so, we not only collected a sizable amount of real-life data, but our consideration of extensive longitudinal data may present new inconsistencies in how participants perceive and cognitively aggregate their past interactions with other groups — which may suggest large-scale recall-biases or conflations of within and between participant effects in conventional cross-section studies.

% Motivation and core needs 
A broader theoretical implication relates to role of situational motivation in intergroup contacts. Our results offer a first promising test of psychological needs in intergroup contact. While our results are tentative given their novelty within the field, they were highly consistent across studies and may offer new theoretical avenues. Psychological needs are a facet of the human experience that has thus far been under–emphasized in the intergroup contact literature. This stands in stark contrast of the many cognitive \citep[e.g.,][]{Pettigrew1998, Brown2005} and emotional aspects investigated within the field \citep[e.g.,][]{Stephan2008, Paolini2004}. However, if motivational mechanisms are indeed shown to be meaningful factors in intergroup contacts, future research may be able to integrate broader theoretical frameworks of intergroup contact (e.g., motivations guiding cognition and affect, which in turn drive behavior. cf., theory of reasoned goal pursuit; \citealp{Ajzen2019}).

% practical use of needs
Additionally, situational motivations in intergroup contact also offer promising avenues for practitioners and policy-makers. Intergroup contact theory is among the most implemented psychological theories \citep[e.g.,][]{Pettigrew2006, AlRamiah2012a, Reimer2021}. Given our findings that psychological needs in every-day intergroup contacts were at least as powerful as Allport's conditions in predicting outgroup attitudes, considerations of people's needs offer a substantially more immediate mechanism to address. In cases where some or all optimal contact conditions are not possible to be fulfilled, needs offer an even more compelling alternative (e.g., where equal status is contextually not possible or in cases where people help despite a lack of institutional support). Additionally, our conceptualization of situational needs might offer an additional opportunity for practitioners and interventions. Instead of addressing needs as a one-size fits all solution (e.g., simply focusing on competence needs), one may at times ask outgroup interaction partners what they need during an interaction. This is not to say that we should not explore which needs tend to be relevant to specific groups in specific intergroup contact contexts. Rather, interventions for which such data is not available or infeasible to collect, a flexible and reactive approach to intergroup contact needs might be more fruitful.

\subsection{Conclusion}
% Conclusion paragraph
In sum, in this paper we used extensive longitudinal methodologies to capture real-life interactions of recent migrants with the majority outgroup. Our three studies showcase the feasibility and utility of such data to test intergroup contact theory. We provide evidence that the fulfillment of situational psychological needs during real-life intergroup contacts meaningfully predicts perceived interaction quality and positive outgroup attitudes. Our results point to motivation as an understudied aspect of intergroup contact that is important in understanding when and why an interaction is perceived as positive and will lead to more positive outgroup attitudes.

>>>>

We have no known conflict of interest to declare. The authors received no specific funding for this work. Materials and  software is available at \url{https://janniscodes.github.io/intergroup-contact-needs/}  \citep{Kreienkamp2022}. Protocols, materials, data, and code are available at \url{https://osf.io/pr9zs/?view_only=208a53a1f0ff48dda1c17357328fa578} \citep{Kreienkamp2022a}. The preregistration of Study 3 can be accessed as part of our Open Science Framework repository \citep{Kreienkamp2021f}

@@

% Abstract
\abstract{
One challenge of modern intergroup contact research has been the question of when and why an interaction is perceived as positive and leads to better intergroup relations. This question becomes especially relevant when considering that negative intergroup contacts are common during everyday life. To understand positive intergroup interactions, we propose to consider situational motivations and psychological needs during everyday intergroup contact. We conducted three extensive longitudinal studies with recent migrants, to capture their interactions with the majority outgroup (total N of measurements = 10,297). Across the three studies, we find inconsistent results for a traditional test of the contact hypothesis with experience sampling data. However, when using multilevel models to test the contact hypothesis, we find that individual outgroup interactions and perceived interaction quality have a strong influence on positive outgroup attitudes. The fulfillment of psychological needs is a strong predictor of positive attitudes after interactions and that these positive effects emerged via perceived interaction quality. The situational needs remained the core predictor of outgroup attitudes even after controlling for Allport’s (\citeyear{Allport1954b}) contact conditions or more general fundamental needs. As one of the first studies to test intergroup contact theory using extensive longitudinal data, we offer insight into the mechanisms of positive intergroup contact during real-life interactions and find situational motivations to be a key building block of understanding and addressing positive intergroup interactions.

@@

\section{Need Mechanism in Intergroup Contact}
% Motivational mechanism:
% interaction quality is important
% past research has either focused on conditions or cognitive-affective processes
% neither are necessary or explain why an interaction is positive
Several meta-analytic reviews have found that positive intergroup contact reduces prejudice and increases positive attitudes in experimental and cross-sectional studies \citep[][]{Tropp2005, Pettigrew2006, Davies2011}, as well as in intergroup contact interventions outside the lab \citep[][]{Beelmann2014, Lemmer2015}. It is widely accepted that equal status, common goals, collaboration, and structural support during the interaction form the optimal conditions for such positive contact effects \citep[Allport's Optimal Contact conditions;][]{Allport1954b, Pettigrew1969}. However, a major meta-analysis of these contact conditions showed that contact resulted in more positive intergroup relations even when Allport's conditions were not met \citep[][]{Pettigrew2006}. It, thus, remains unclear when and why exactly an interaction is perceived as positive and what factors (beyond Allport's conditions) explain positive contact effects\footnote{It should be noted here that following Allport's original conditions, several additional conditions of optimal contact were proposed \citep[for a critical discussion see][]{Pettigrew1986}. Similarly, since the early 21\textsuperscript{st} century many psychological processes during intergroup contact were examined \citep[e.g. see,][]{Paolini2021}. Among others, researchers have for example explored different forms of social categorizations \citep[][]{Pettigrew1998}, the salience of social categories \citep[][]{Brown2005}, intimacy \citep[e.g.,][]{Marinucci2021} and attachment \citep[e.g.,][]{Tropp2021}, threat and intergroup anxiety \citep[e.g.,][]{Stephan2008, Paolini2004}, and to a lesser extent knowledge about the other group \citep[][]{Pettigrew2008c}. However, all these advances share the underlying criticism that they are not geared towards understanding when and why an interaction is perceived as positive.}. We propose that we may look to the literature on motivation and psychological needs to understand when and why exactly an interaction is perceived as positive.

@@

\section{Real-Life Process Data}
While we have argued that a need fulfillment mechanism is relevant to intergroup contact generally, its flexible and broad applicability might be ideally suited to address the pressing issue of understanding natural intergroup contacts outside the lab. Investigations of such `real-life interactions' often suffer from the difficulty that past intergroup contact research has either focused on the mechanisms of individual interactions in artificial lab studies or has focused on longer-term recall self-reports of natural interactions \citep[e.g.,][]{Pettigrew2006}. Even with extended intervention studies, the most fine-grained data available is usually limited to pre-post-control designs. Such empirical practices stand, however, in stark contrast to many of the theoretical advances that have focused on the dynamic nature of intergroup relations \citep[e.g.,][]{Pettigrew1998}, as well as the core idea of the contact hypothesis, which was focused on the daily interactions of people \citep[see][]{Allport1954b}. As a result, prominent figures within the intergroup contact field have called for studies that collect longitudinal \citep[][]{Pettigrew1998, Pettigrew2008, Pettigrew2008b, Pettigrew2011} and real-life experience-sampling data outside the lab \citep[][]{MacInnis2015, McKeown2017, Dixon2005}. Such data would be able to capture real-life interactions that include interaction-specific mechanism information close the actual experience\footnote{Additionally, such experience-sampling data can be collected close to the intergroup interactions and would, thus, largely mitigate recall biases. Moreover, because data is nested within participants, experience-sampling data often allows to capture large amounts of high-quality data with relatively few participants.}.

@@

To test our main hypotheses, we collected three independent sets of extensive longitudinal data (Studies 1–3). The full surveys are available in Supplementary Materials A as well as in our open science repository \citep[including a complete codebook, see][]{KreienkampMasked2022a}. Our most comprehensive study (Study 3) was fully preregistered  \citep[available at][]{KreienkampMasked2021f}. All studies received ethical approval from the University Masked for Peer Review. The fully annotated analyses are available in Supplementary Material B. As is common with multilevel analyses, we use a hierarchical modeling approach and always report the final model in-text (for the full modeling process see Supplementary Material B).

@@

\section{Discussion}
% aims re-iterated
The aim of this project was threefold. We (1) tested the basic predictions of the intergroup contact hypothesis and Allport's optimal conditions in real-life extensive longitudinal data. As a flexible alternative, we (2) proposed that the fulfillment of situational psychological needs meaningfully predicts positive outgroup attitudes. Finally, we (3) assessed whether the fulfillment of psychological needs (and Allport's conditions) have their effect (partially) through perceived interaction quality.

% Contact hypothesis in ESM data [inconsistent in aggregate but consistent in ML analysis]
When considering the results of the three studies jointly, we found mixed results for the basic intergroup contact hypothesis. More specifically, we observe that separating between-participant and within-participant effects is crucial. For between-subjects effects (using aggregated reports), neither the number of interactions, nor the average interaction quality were significant predictors of positive outgroup attitudes across all studies (also see Figure \ref{fig:ContactHypothesis}a). These aggregate findings are surprising because most past cross-sectional studies that have investigated the effect of interaction frequency have also considered an aggregate measurement (albeit not one from actual daily reports). The absence of this aggregate effect could underline the fact that cross-sectional retrospective data might be misleading because (1) it presents a mixture of within and between-subjects effects \citep[][]{Hamaker2020}, or (2) suffers from recall biases (e.g., where times with no outgroup interactions are undervalued by participants during the retrospective evaluations).\footnote{It should be noted that such inconsistencies with past research might in part be a data artefact (e.g., because most people reported substantially more measurements during which they did not have an outgroup interaction). However, the models we employed should in principle be resistant to such imbalances.} In support of this possibility, once we used multilevel modeling to separate effects into a between-subjects and within-subjects part, we found that when having an interaction with an outgroup member (vs. not having an outgroup interaction) is a strong predictor of outgroup attitudes across all three studies (even when taking other non-outgroup interactions into account; also see Figure \ref{fig:ContactHypothesis}b). We thus find that when we aggregate interaction frequencies and interaction quality from our intensive longitudinal data, we find inconsistent results from the past literature, however within participants, intergroup contacts are significantly related with positive outgroup attitudes.

@@

\subsection{Limitations}
% (1) Sample = minority and (voluntary) migrants, (2) Methodology = short scales (reliability) and no lagged effects, (3) Core Need Concept = non-specific
Among the limitations of our studies we should consider the following. One limitation is our choice of samples. While we believe that a need–based mechanism should be relevant to any inter-group contact, our sample has focused on a minority- and (voluntary) migrant perspective. Without additional evidence it thus remains difficult to judge whether motivational effects will generalize to other migrant groups (e.g., forced migrants), other intergroup contexts (e.g., gender-, religious-, or sexual orientation groups), and to majority groups in their outgroup attitudes. We know of no research that for other contexts psychological needs would not be relevant (once existential needs are met) but future research may extend our findings to other contexts to build an even broader understanding of psychological needs in intergroup contacts. A second limitation lies in our methodology. While extensive longitudinal data is close to real-life events, this method comes at the expense of longer and more robust scales. Long and repetitive scales are often not feasible in extensive longitudinal methods because of the increased burden to the participants. To circumvent this shortcoming, we have ensured that the measures we used were, whenever possible, based on past validations. However, the circumstance remains that extensive longitudinal data often does not allow the same scrutiny of measurement reliability as single–shot cross–section data sets. An additional methodological question lies in the unexplored potential of the longitudinal aspects of our data. For our research questions, we have focused on contemporaneous effects within the data set, yet future investigations should seek to extend the mechanism to developmental trajectories within and between participants. And finally, our conceptualization of core situational needs has been focused on the most essential test of a motivational mechanism. This comes at the expense of specificity in the situational needs (i.e., we have not explored which exact needs people had during specific interactions). With the adaptive measurement we used such an investigation would be possible (e.g., by clustering or classifying the free text entries) but would not have been relevant to our theory–focused research question. Nonetheless, future research may explore which exact needs are most important in different intergroup contexts.

\subsection{Implications}
% ESM data
Given these limitations, we can nonetheless draw a number of implications for other researchers and practitioners --- ranging from the benefits of longitudinal data to theoretical implications. A first implication concerns the feasibility and usefulness of extensive longitudinal data for intergroup contact research and the broader field of social psychology. While setting up an extensive longitudinal study is not easy, we believe the efforts to be similar to a sizable cross-sectional data collection (i.e., for a longitudinal or a high-quality cross-sectional data set with over 10,000 data-points). Given the comparable effort, this opens up the possibility to explore research questions that focus on real-life phenomena outside the lab or focus on phenomena that depend on changes and influences over time. In the context of intergroup contact research, we were among the first to answer calls to test intergroup contact mechanisms using on extended real-life data \citep[e.g.,][]{Pettigrew2011, MacInnis2015}. In doing so, we not only collected a sizable amount of real-life data, but our consideration of extensive longitudinal data may present new inconsistencies in how participants perceive and cognitively aggregate their past interactions with other groups — which may suggest large-scale recall-biases or conflations of within and between participant effects in conventional cross-section studies.

% Motivation and core needs 
A broader theoretical implication relates to role of situational motivation in intergroup contacts. Our results offer a first promising test of psychological needs in intergroup contact. While our results are tentative given their novelty within the field, they were highly consistent across studies and may offer new theoretical avenues. Psychological needs are a facet of the human experience that has thus far been under–emphasized in the intergroup contact literature. This stands in stark contrast of the many cognitive \citep[e.g.,][]{Pettigrew1998, Brown2005} and emotional aspects investigated within the field \citep[e.g.,][]{Stephan2008, Paolini2004}. However, if motivational mechanisms are indeed shown to be meaningful factors in intergroup contacts, future research may be able to integrate broader theoretical frameworks of intergroup contact (e.g., motivations guiding cognition and affect, which in turn drive behavior. cf., theory of reasoned goal pursuit; \citealp{Ajzen2019}. Also see \citealp{Kreienkamp2022d}).

% practical use of needs
Additionally, situational motivations in intergroup contact also offer promising avenues for practitioners and policy-makers. Intergroup contact theory is among the most implemented psychological theories \citep[e.g.,][]{Pettigrew2006, AlRamiah2012a, Reimer2021}. Given our findings that psychological needs in every-day intergroup contacts were at least as powerful as Allport's conditions in predicting outgroup attitudes, considerations of people's needs offer a substantially more immediate mechanism to address. In cases where some or all optimal contact conditions are not possible to be fulfilled, needs offer an even more compelling alternative (e.g., where equal status is contextually not possible or in cases where people help despite a lack of institutional support). Additionally, our conceptualization of situational needs might offer an opportunity for practitioners and interventions. Instead of addressing needs as a one-size fits all solution (e.g., simply focusing on competence needs), one may at times ask outgroup interaction partners what they need during an interaction. This is not to say that we should not explore which needs tend to be relevant to specific groups in specific intergroup contact contexts. Rather, during interventions for which data on important need contents are not available or infeasible to collect, a flexible and reactive approach of inquiring momentary intergroup contact needs might be more fruitful.

\subsection{Conclusion}
% Conclusion paragraph
In sum, we used extensive longitudinal methodologies to capture real-life interactions of recent migrants with the majority outgroup. Our three studies showcase the feasibility and utility of such data to test intergroup contact theory. We provide evidence that the fulfillment of situational psychological needs during real-life intergroup contacts meaningfully predicts perceived interaction quality and positive outgroup attitudes. Our results point to motivation as an understudied aspect of intergroup contact that is important in understanding when and why an interaction is perceived as positive and will lead to more positive outgroup attitudes.
